FinalManuscript_DISCS_UG

### 2019-03-03 parts/00_abstract.tex
Misc edits (Ch2,3,appendix,abstract, acknowledgements)

\begin{thesisabstract}
    \noindent
    This paper is a sample document that serves as a format and content guideline for undergraduate thesis submissions to the Department of Information Systems and Computer Science. In this section, the abstract, the group should be able to give the readers a clear and concise overview of their study. The section should contain the objectives of the thesis, the methods to be used, and when available, the results of the study, the conclusion, and the recommendations for further work, all based on the intended research objectives. A good abstract should be at most around 150--200 words, or half a page. It should also not contain any references, figures, or equations.
\end{thesisabstract}

>>>>

\begin{thesisabstract}
    \noindent
    Homomorphic cryptography allows for encrypted data to be modified and operated on without requiring decryption, although current homomorphic cryptosystems are either limited in their permitted operations or are significantly more time-intensive than standard non-homomorphic cryptosystems. Regardless, homomorphic cryptography has been targeted for use in secure image processing and facial recognition, due to their ability to maintain data privacy. In this paper, we compared the viability of the Paillier, Damg{\aa}rd--Geisler--Kr{\o}igaard (DGK), and Brakerski--Gentry--Vaikuntanathan (BGV) cryptosystems for facial image processing applications, by implementing a software library equipped with these cryptosystems, and comparing their time efficiency and accuracy. Furthermore, as an extension to previous research, we attempt to support non-linear image processing operations by deriving and implementing applicable closed-form approximations. Preliminary results have shown that the Paillier and DGK cryptosystems are comparable in accuracy and may be used for image negation, but only the Paillier cryptosystem is consistent enough to produce reasonable to accurate results.

    % This paper is a sample document that serves as a format and content guideline for undergraduate thesis submissions to the Department of Information Systems and Computer Science. In this section, the abstract, the group should be able to give the readers a clear and concise overview of their study. The section should contain the objectives of the thesis, the methods to be used, and when available, the results of the study, the conclusion, and the recommendations for further work, all based on the intended research objectives. A good abstract should be at most around 150--200 words, or half a page. It should also not contain any references, figures, or equations.
\end{thesisabstract}

### 2019-03-03 parts/02_relatedliterature_05partiallyhomomorphic.tex
Misc edits (Ch2,3,appendix,abstract, acknowledgements)

\subsubsection{Extension to Floating-Point Numbers}
\label{sec:fp_operations}
We can use the following protocol described in~\cite{ziad_cryptoimg:_2016} in order to extend Paillier and DGK to floating-point numbers.

>>>>

\subsubsection{Extension to Floating-Point Numbers}
\label{sec:fp_number_extension}
We can use the following protocol described in~\cite{ziad_cryptoimg:_2016} in order to extend Paillier and DGK to floating-point numbers.

### 2019-03-03 parts/02_relatedliterature_06fullyhomomorphic.tex
Misc edits (Ch2,3,appendix,abstract, acknowledgements)

\subsection{DGHV Cryptosystem}
In 2009, Dijk, Gentry, Halevi, and Vaikuntanathan created a fully homomorphic cryptosystem which operates primarily through elementary modular arithmetic, commonly called the DGHV cryptosystem \cite{cryptoeprint:2009:616}. This was further improved in 2011 by Coron, Naccache, and Tibouchi, \cite{cryptoeprint:2011:277, cryptoeprint:2011:440} to support shorter public keys, reducing the size of secure public keys from 12.5 GB to 2.5 MB.

>>>>

\subsection{The DGHV Cryptosystem}
In 2009, Dijk, Gentry, Halevi, and Vaikuntanathan created a fully homomorphic cryptosystem which operates primarily through elementary modular arithmetic, commonly called the DGHV cryptosystem \cite{cryptoeprint:2009:616}. This was further improved in 2011 by Coron, Naccache, and Tibouchi, \cite{cryptoeprint:2011:277, cryptoeprint:2011:440} to support shorter public keys, reducing the size of secure public keys from 12.5 GB to 2.5 MB.

### 2019-03-03 parts/08_appendices_b_numericalapproximations_updated.tex
Misc edits (Ch2,3,appendix,abstract, acknowledgements)

\subsection{Known Approximations for Transcendental Functions}
Many methods to compute transcendental functions implemented in computer hardware and software libraries rely on direct knowledge of the function arguments. Several such approaches include CORDIC (coordinate rotation digital computer), pseudo-division and pseudo-multiplication algorithms~\cite{walther_cordic_2000}, as well as minimax polynomials, which are determined numerically using the Remez algorithm~\cite{harrison_computation_1999}. These algorithms, while efficient, rely on table lookups, which cannot be done in a privacy-preserving system where the function arguments cannot be directly determined.

@@

Khattri presents the following method to approximate the function $f(x)=\log\left(1+x\right)$.
Let $x = 1/n$ and consider the integral
\begin{align}
	\label{eq:log_integral}
	\int_{n}^{n+1}{\frac{1}{t}\diff t}=\log{\left(1+\frac{1}{n}\right)}.
\end{align}

@@

While the closed-form approximation presented by Khattri (Equation~\ref{eq:standard_logarithm_quadrature}) is accurate for values of $x$ near zero, it diverges from $\log{\left(1+x\right)}$ significantly for large values of $x$, as shown in Figure~\ref{fig:standard_logarithm_quadrature}.
\begin{figure}[!ht]
	\centering
	\begin{tikzpicture}
		\begin{axis}[
			% height=6cm, width=9cm,
			xmin=0, xmax=300,
			ymin=0, ymax=5.5, ytick={0,...,6},
			xlabel={$x$}, ylabel={$y$},
			grid=major,
			legend pos=south east,
			legend cell align=left,
		]
		\addplot[very thick, ACMDarkBlue, domain=0:300, samples=100]{1/30*(137*x^5 + 2310*x^4 + 9870*x^3 + 15120*x^2 + 7560*x)/(x^5 + 30*x^4 + 210*x^3 + 560*x^2 + 630*x + 252)};
		\addplot[densely dashed, domain=0:300, samples=100]{ln(x)};
		\legend{5-point quadrature approximation, Actual value of $\log(1+x)$}
		\end{axis}
	\end{tikzpicture}%
	\caption{Graph of $\log{\left(1+x\right)}$ and the approximation in Equation~\ref{eq:standard_logarithm_quadrature}}
	\label{fig:standard_logarithm_quadrature}
\end{figure}

@@

\section{Logarithmic Function Approximation}
\label{sec:logarithm_approximation}
We now present our main contribution in this paper: a general method to improve the general accuracy of the logarithm approximation given by Khattri in~\cite{khattri_new_2009} for use in privacy-preserving calculation, provided that accuracy is only required for a specific range.

For this approximation, we assume we only require accurate results in the input range $x \in [0, 255]$. This reflects some potential applications such as in logarithm intensity transformations in image processing~\cite{gonzalez_digital_2008}.

@@

We note that the approximation in Equation~\ref{eq:optimal_log_approximation} requires the value of $\log{16}$. This can be computed in the plaintext domain and added to an encrypted value using the properties of the Paillier cryptosystem. The remaining arithmetic operations in the approximation can be computed in the encrypted domain, using the protocols described in Section~\ref{sec:fp_arithmetic}. Therefore, this approximation can be calculated in a privacy-preserving manner.

@@

\begin{figure}[ht]
	\centering
    \begin{tikzpicture}
		\begin{semilogyaxis}[
			% height=6cm, width=9cm,
			xmin=0, xmax=300,
			ymin=1e-10, ymax=1.4,
			xlabel={$x$}, ylabel={Absolute error},
			grid=major,
			legend pos=south east,
			legend cell align=left,
			tick scale binop=\times,
		]
		% \addplot[very thick, domain=0:300, samples=500]{abs(-1/30*(x^5*(30*ln(1/16) - 137) + 45*x^4*(270*ln(1/16) - 593) + 30*x^3*(27210*ln(1/16) - 20579) + 30*x^2*(488810*ln(1/16) + 216621) + 45*x*(1690670*ln(1/16) + 2694207) + 93677310*ln(1/16) + 257804775)/(x^5 + 405*x^4 + 27210*x^3 + 488810*x^2 + 2536005*x + 3122577) - ln(x + 1))};
		% \addplot[very thick, domain=0:300, samples=300]{abs(1/30*(137*x^5 + 26685*x^4 + 617370*x^3 - 6498630*x^2 - 121239315*x - 257804775)/(x^5 + 405*x^4 + 27210*x^3 + 488810*x^2 + 2536005*x + 3122577) - ln(1/16) - ln(x + 1))};
		\addplot[very thick, ACMDarkBlue] table[x=x, y=error] {plots/abs_err_scaledlog.txt};
		\addplot[densely dashed, domain=0:300, samples=100]{abs(1/30*(137*x^5 + 2310*x^4 + 9870*x^3 + 15120*x^2 + 7560*x)/(x^5 + 30*x^4 + 210*x^3 + 560*x^2 + 630*x + 252) - ln(x + 1))};
		\legend{Scaled quadrature ($\alpha = 1/16$), Original quadrature ($\alpha = 1$)}
	\end{semilogyaxis}
	\end{tikzpicture}%
	\caption{Comparison of the absolute error between the original approximation in Equation~\ref{eq:standard_logarithm_quadrature} and the scaled approximation in Equation~\ref{eq:optimal_log_approximation}}
	\label{fig:log_error_comparison}
\end{figure}

@@

Let $I\left(x,\alpha\right)$ be the function obtained by approximating $\int_{n}^{\alpha n + \alpha}{\frac{1}{x}\diff x}$ with five-point Gauss--Legendre quadrature. The absolute error of the approximation compared to $\log{\left(1+x\right)}$ at a given $x$ and $\alpha$ is given by the following function.
\begin{align*}
	A\left(x,\alpha\right) = \left| I\left(x,\alpha\right) - \log\alpha - \log{\left(1+x\right)}\right|.
\end{align*}
We want to show that the function
\begin{align*}
	f\left(\alpha\right) = \max_{0\leq x \leq 255}{A\left(x,\alpha\right)}
\end{align*}
is minimized at $\alpha = 1/16$.
We consider the functions $A\left(0,\alpha\right)$ and $A\left(255,\alpha\right)$. These are plotted in Figure~\ref{fig:endpoint_plot}.
\begin{figure}[!ht]
	\centering
    \begin{tikzpicture}
		\begin{semilogyaxis}[
			% height=6cm, width=9cm,
			xmin=0, xmax=1,
			ymin=5e-7, ymax=1.3,
			xlabel={$\alpha$}, ylabel={Absolute error},
			grid=major,
			extra x ticks={1/16},
			extra x tick labels={$\frac{1}{16}$},
			legend pos=south east,
			legend cell align=left,
			tick scale binop=\times,
		]
		\addplot[thick, ACMDarkBlue, domain=0:1] table[x=alpha, y=x0] {plots/abs_err_sl_0_255.txt};
		\addplot[thick, ACMOrange, domain=0:1] table[x=alpha, y=x255] {plots/abs_err_sl_0_255.txt};
		\legend{$x=0$, $x=255$}
		% \legend{$A\left(0,\alpha \right)$, $A\left(255,\alpha \right)$}
	\end{semilogyaxis}
	\end{tikzpicture}%
	\caption{Graphs of $A\left(x,\alpha\right)$ at $x=0$ and $x=255$}
	\label{fig:endpoint_plot}
\end{figure}

@@

We then consider the function $A\left(x,1/16\right)$, shown in Figure~\ref{fig:single_alpha_plot}.
It was determined numerically that the local maxima of $A\left(x,1/16\right)$ on the interval occur at $x=0$ and $x=255$. Since a scaling factor other than $\alpha=1/16$ yields a higher absolute error at either $x=0$ or $x=255$, as seen in Figure~\ref{fig:endpoint_plot}, it follows that a scaling factor of $\alpha = 1/16$ minimizes the maximum absolute error of the scaled approximation for $\log{\left(1+x\right)}$, within the range $x \in [0, 255]$.
\begin{figure}[ht]
	\centering
	\begin{tikzpicture}
		\begin{semilogyaxis}[
			xmin=0, xmax=300,
			ymin=1e-15, ymax=0.035,
			xlabel={$x$}, ylabel={Absolute error},
			grid=major,
			legend pos=south east,
			legend cell align=left,
			tick scale binop=\times,
		]
		% \addplot[very thick, domain=0:300, samples=300]{abs(1/30*(137*x^5 + 26685*x^4 + 617370*x^3 - 6498630*x^2 - 121239315*x - 257804775)/(x^5 + 405*x^4 + 27210*x^3 + 488810*x^2 + 2536005*x + 3122577) - ln(1/16) - ln(x + 1))};
		\addplot[thick, ACMDarkBlue] table[x=x, y=error] {plots/abs_err_scaledlog.txt};
	\end{semilogyaxis}
	\end{tikzpicture}%
	\caption{Graph of $A\left(x,1/16\right)$, the absolute error of the approximation in Equation~\ref{eq:optimal_log_approximation} given a scaling factor of $\alpha=1/16$}
	\label{fig:single_alpha_plot}
\end{figure}

We have shown how we can adapt Khattri's approximation to allow for privacy-preserving logarithm computation over a larger interval. We note we only require the quadrature approach if series representations of the desired function are not suitable for privacy-preserving calculation using the floating-point arithmetic supported by the Paillier cryptosystem.

As a concrete example, evaluating partial sums of the series representation for the exponential function,
\begin{align*}
	e^x = \sum_{n=0}^{\infty}{\frac{x^n}{n!}},
\end{align*}
is a valid approach for privacy-preserving computation, as this series converges for all $x$.

>>>>

\chapter{Logarithm Approximation}
Many methods to compute transcendental functions implemented in computer hardware and software libraries rely on direct knowledge of the function arguments. Several such approaches include CORDIC (coordinate rotation digital computer), pseudo-division and pseudo-multiplication algorithms~\cite{walther_cordic_2000}, as well as minimax polynomials, which are determined numerically using the Remez algorithm~\cite{harrison_computation_1999}. These algorithms, while efficient, rely on table lookups, which cannot be done in a privacy-preserving system where the function arguments cannot be directly determined.

@@

\section{Known Closed-Form Logarithm Approximations}
Khattri ~\cite{khattri_new_2009} presents the following method to approximate the function $f(x)=\log\left(1+x\right)$.
Let $x = 1/n$ and consider the integral
\begin{align}
	\label{eq:log_integral}
	\int_{n}^{n+1}{\frac{1}{t}\diff t}=\log{\left(1+\frac{1}{n}\right)}.
\end{align}

@@

While the closed-form approximation presented by Khattri (Equation~\ref{eq:standard_logarithm_quadrature}) is accurate for values of $x$ near zero, it diverges from $\log{\left(1+x\right)}$ significantly for large values of $x$, as shown in Figure~\ref{fig:standard_logarithm_quadrature}.
\begin{figure}[!ht]
	\centering
	% \begin{tikzpicture}
	% 	\begin{axis}[
	% 		% height=6cm, width=9cm,
	% 		xmin=0, xmax=300,
	% 		ymin=0, ymax=5.5, ytick={0,...,6},
	% 		xlabel={$x$}, ylabel={$y$},
	% 		grid=major,
	% 		legend pos=south east,
	% 		legend cell align=left,
	% 	]
	% 	\addplot[very thick, blue, domain=0:300, samples=100]{1/30*(137*x^5 + 2310*x^4 + 9870*x^3 + 15120*x^2 + 7560*x)/(x^5 + 30*x^4 + 210*x^3 + 560*x^2 + 630*x + 252)};
	% 	\addplot[densely dashed, domain=0:300, samples=100]{ln(x)};
	% 	\legend{5-point quadrature approximation, Actual value of $\log(1+x)$}
	% 	\end{axis}
	% \end{tikzpicture}%
	\caption{Graph of $\log{\left(1+x\right)}$ and the approximation in Equation~\ref{eq:standard_logarithm_quadrature}}
	\label{fig:standard_logarithm_quadrature}
\end{figure}

@@

\section{Scaled Approximation}
\label{sec:logarithm_approximation}
This appendix a general method to improve the general accuracy of the logarithm approximation given by Khattri in~\cite{khattri_new_2009} for use in privacy-preserving calculation, provided that accuracy is only required for a specific range.

For this approximation, we assume we only require accurate results in the input range $x \in [0, 255]$, to reflect use as part of the logarithm image transformation.

@@

We note that the approximation in Equation~\ref{eq:optimal_log_approximation} requires the value of $\log{16}$. This can be computed in the plaintext domain and added to an encrypted value using the properties of the Paillier or DGK cryptosystems, while an FHE scheme would have to encrypt the value of $\log{16}$. The remaining arithmetic operations in the approximation can be computed in the encrypted domain, using the protocols described in Section~\ref{sec:fp_arithmetic}. Therefore, this approximation can be calculated in a privacy-preserving manner.

@@

\begin{figure}[ht]
	\centering
  %   \begin{tikzpicture}
	% 	\begin{semilogyaxis}[
	% 		% height=6cm, width=9cm,
	% 		xmin=0, xmax=300,
	% 		ymin=1e-10, ymax=1.4,
	% 		xlabel={$x$}, ylabel={Absolute error},
	% 		grid=major,
	% 		legend pos=south east,
	% 		legend cell align=left,
	% 		tick scale binop=\times,
	% 	]
	% 	% \addplot[very thick, domain=0:300, samples=500]{abs(-1/30*(x^5*(30*ln(1/16) - 137) + 45*x^4*(270*ln(1/16) - 593) + 30*x^3*(27210*ln(1/16) - 20579) + 30*x^2*(488810*ln(1/16) + 216621) + 45*x*(1690670*ln(1/16) + 2694207) + 93677310*ln(1/16) + 257804775)/(x^5 + 405*x^4 + 27210*x^3 + 488810*x^2 + 2536005*x + 3122577) - ln(x + 1))};
	% 	% \addplot[very thick, domain=0:300, samples=300]{abs(1/30*(137*x^5 + 26685*x^4 + 617370*x^3 - 6498630*x^2 - 121239315*x - 257804775)/(x^5 + 405*x^4 + 27210*x^3 + 488810*x^2 + 2536005*x + 3122577) - ln(1/16) - ln(x + 1))};
	% 	\addplot[very thick, blue] table[x=x, y=error] {plots/abs_err_scaledlog.txt};
	% 	\addplot[densely dashed, domain=0:300, samples=100]{abs(1/30*(137*x^5 + 2310*x^4 + 9870*x^3 + 15120*x^2 + 7560*x)/(x^5 + 30*x^4 + 210*x^3 + 560*x^2 + 630*x + 252) - ln(x + 1))};
	% 	\legend{Scaled quadrature ($\alpha = 1/16$), Original quadrature ($\alpha = 1$)}
	% \end{semilogyaxis}
	% \end{tikzpicture}%
	\caption{Comparison of the absolute error between the original approximation in Equation~\ref{eq:standard_logarithm_quadrature} and the scaled approximation in Equation~\ref{eq:optimal_log_approximation}}
	\label{fig:log_error_comparison}
\end{figure}

@@

Let $I\left(x,\alpha\right)$ be the function obtained by approximating $\int_{n}^{\alpha n + \alpha}{\frac{1}{x}\diff x}$ with five-point Gauss--Legendre quadrature. The absolute error of the approximation compared to $\log{\left(1+x\right)}$ at a given $x$ and $\alpha$ is given by the following function.
\begin{align*}
	A\left(x,\alpha\right) = \left| I\left(x,\alpha\right) - \log\alpha - \log{\left(1+x\right)}\right|.
\end{align*}
We want to show that the function
\begin{align*}
	f\left(\alpha\right) = \max_{0\leq x \leq 255}{A\left(x,\alpha\right)}
\end{align*}
is minimized at $\alpha = 1/16$.
We consider the functions $A\left(0,\alpha\right)$ and $A\left(255,\alpha\right)$. These are plotted in Figure~\ref{fig:endpoint_plot}.
\begin{figure}[!ht]
	\centering
  %   \begin{tikzpicture}
	% 	\begin{semilogyaxis}[
	% 		% height=6cm, width=9cm,
	% 		xmin=0, xmax=1,
	% 		ymin=5e-7, ymax=1.3,
	% 		xlabel={$\alpha$}, ylabel={Absolute error},
	% 		grid=major,
	% 		extra x ticks={1/16},
	% 		extra x tick labels={$\frac{1}{16}$},
	% 		legend pos=south east,
	% 		legend cell align=left,
	% 		tick scale binop=\times,
	% 	]
	% 	\addplot[thick, ACMDarkBlue, domain=0:1] table[x=alpha, y=x0] {plots/abs_err_sl_0_255.txt};
	% 	\addplot[thick, ACMOrange, domain=0:1] table[x=alpha, y=x255] {plots/abs_err_sl_0_255.txt};
	% 	\legend{$x=0$, $x=255$}
	% 	% \legend{$A\left(0,\alpha \right)$, $A\left(255,\alpha \right)$}
	% \end{semilogyaxis}
	% \end{tikzpicture}%
	\caption{Graphs of $A\left(x,\alpha\right)$ at $x=0$ and $x=255$}
	\label{fig:endpoint_plot}
\end{figure}

@@

We then consider the function $A\left(x,1/16\right)$, shown in Figure~\ref{fig:single_alpha_plot}.
It was determined numerically that the local maxima of $A\left(x,1/16\right)$ on the interval occur at $x=0$ and $x=255$. Since a scaling factor other than $\alpha=1/16$ yields a higher absolute error at either $x=0$ or $x=255$, as seen in Figure~\ref{fig:endpoint_plot}, it follows that a scaling factor of $\alpha = 1/16$ minimizes the maximum absolute error of the scaled approximation for $\log{\left(1+x\right)}$, within the range $x \in [0, 255]$.
\begin{figure}[ht]
	\centering
	% \begin{tikzpicture}
	% 	\begin{semilogyaxis}[
	% 		xmin=0, xmax=300,
	% 		ymin=1e-15, ymax=0.035,
	% 		xlabel={$x$}, ylabel={Absolute error},
	% 		grid=major,
	% 		legend pos=south east,
	% 		legend cell align=left,
	% 		tick scale binop=\times,
	% 	]
	% 	% \addplot[very thick, domain=0:300, samples=300]{abs(1/30*(137*x^5 + 26685*x^4 + 617370*x^3 - 6498630*x^2 - 121239315*x - 257804775)/(x^5 + 405*x^4 + 27210*x^3 + 488810*x^2 + 2536005*x + 3122577) - ln(1/16) - ln(x + 1))};
	% 	\addplot[thick, ACMDarkBlue] table[x=x, y=error] {plots/abs_err_scaledlog.txt};
	% \end{semilogyaxis}
	% \end{tikzpicture}%
	\caption{Graph of $A\left(x,1/16\right)$, the absolute error of the approximation in Equation~\ref{eq:optimal_log_approximation} given a scaling factor of $\alpha=1/16$}
	\label{fig:single_alpha_plot}
\end{figure}

We have shown how we can adapt Khattri's approximation to allow for privacy-preserving logarithm computation over a larger interval.

\section{Approximation for $x^\gamma$}
To approximate $x^\gamma$ for any $\gamma \in \mathbb{R}$, we rewrite $x^\gamma$ as follows:
\begin{align*}
  x^\gamma = e^{\log{x^\gamma}} = e^{\gamma\log{x}}.
\end{align*}
This expression can then be approximated using the Maclaurin series expansion for $e^x$, which converges for all $x$.
\begin{align*}
  e^x &= \sum_{n=0}^{\infty}{\frac{x^n}{n!}}\\
  \Rightarrow e^{\gamma\log{x}} &= \sum_{n=0}^{\infty}{\frac{(\gamma\log{x})^n}{n!}}
\end{align*}
As we already have an approximation for the natural logarithm, we can evaluate partial sums of the above infinite series to arrive at approximations for $x^\gamma$.
